A quadratic-time algorithm for the cardinality-constrained college application problem

### 2022-07-11 paper_alma/paper.tex
preliminaries, nestedness pf, greedy algorithm

In the present study, we formulate the student's college application decision as the following optimization problem:
\begin{align} \label{headlineproblem}
\begin{split}
\text{maximize}\quad & v(\mathcal{X}) =  \operatorname{E}\Bigl[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~|\mathcal{X}| \leq H
\end{split}
\end{align}
Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $h > 0$ is the number of students the student can apply to; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j\geq 0$ is a utility parameter.

Consider an admissions market with $m$ colleges. The $j$th college is named $c_j$. Consider a single prospective student in this market, and let each $t_j$-value indicate the utility she associates with attending $c_j$, where her utility is $t_0$ if she does not attend college. Let $g_j$ denote the application fee for $c_j$ and $H$ the student's total budget to spend on application fees. Lastly, let $f_j$ denote the student's probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not. It is appropriate to assume that the $Z_j$ are statistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates). Then the student's objective is to maximize the expected utility associated with the best school she is admitted to. Therefore, her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies. 
We focus on the special case in which students are constrained in the \emph{number} of schools to which they can apply, and show that the optimal portfolios in this case are nested in the budget constraint, implying the optimality of the greedy algorithm that adds schools one at a time according to how much they increase the objective value.

@@

The objective function is a nondecreasing submodular function in the sense first described by Nemhauser and Wolsey (1979). They showed that the greedy algorithm is a good approximation; we show that the same algorithm is  exact.

@@

@@

\section{Model} \label{sectionModel}

@@

@@

\parskip 0em
\leftskip 2em
\parindent -2em
Acharya, Mohan S., Asfia Armaan, and Aneeta S. Antony. 2019. ``A Comparison of Regression Models for Prediction of Graduate Admissions.'' In \emph{Second International Conference on Computational Intelligence in Data Science.} \url{https://doi.org/10.1109/ICCIDS.2019.8862140}.

@@

\ifen \subsection{Structure of this paper} \else \subsection{본 논문의 구성}\fi
\ifen Section \ref{preliminaries} introduces some additional notation and assumptions that can be imposed without loss of generality. We also introduce a useful variable-elimination technique and prove that the objective function is submodular.

In Section \ref{homogappcosts}, we consider the special case where each $g_j = 1$ and $H$ is an integer $h \leq m$.  We show that an intuitive heuristic is in fact a $1/h$-approximation algorithm. Then, we show that the optimal portfolios are nested in the budget constraint, which yields an exact algorithm that runs in $O(hm)$ time.

In Section \ref{hetappcosts}, we turn to the scenario in which colleges differ in their application fees. We show that the decision form of the portfolio optimization problem is NP-complete through a polynomial reduction from the binary knapsack problem. We provide four algorithms for this more general setup. The first is a branch-and-bound routine. The second is a dynamic program that iterates on total expenditures and produces an exact solution in pseudopolynomial time, namely $O(Hm + m \log m)$. The third is a different dynamic program that iterates on truncated portfolio valuations. It yields a fully polynomial-time approximation scheme that produces a $(1 - \varepsilon)$-optimal solution in $O(m^3 / \varepsilon)$ time.  The fourth is a simulated-annealing heuristic algorithm that demonstrates strong performance in our synthetic instances.

In Section \ref{numericalexperiments}, we present the results of computational experiments that confirm the validity and time complexity results established in the previous two sections. We also investigate the empirical accuracy of the simulated-annealing heuristic.

In the conclusion, we identify possible improvements to our solution algorithms and introduce a few extensions of the model that capture the features of real-world admissions processes.
\else
\ref{preliminaries}절은 표기법과 일반성을 제한하지 않는 편의적 가정을 제시한다. 유용한 변수 소거 기술도 제시하고 목적함수가 submodularity함을 
증명한다.

\ref{homogappcosts}절에서 각 $g_j = 1$이고 $H$가 $h \leq m$이고 자연수인 특수한 경우를 고려한다. 직관적인 휴리스틱 해법이 실제로 $1/h$-근사 해법임을 증명한다. 그다음에 최적 포트폴리오의 예산 제약에 대한 포함 사슬 관계를 보이고 이를 이용하여 $O(hm)$ 시간 정확한 해법을 도출한다.

\ref{hetappcosts}절에서 각 학교의 지원 비용이 동일하지 않은 일반적인 상황을 다룬다. 이진 배낭 문제에서의 다항 변환을 통하여 포트폴리오 최적화 문제가 NP-complete함을 증명한다. 또한 3개의 일반적인 해법을 제안한다. 첫째는 분지한계법이다. 둘째는 총 지출액으로 탐색하는 $O(Hm + m \log m)$과 같은 의사 다항 시간 동적 계획 해법이다. 셋째는 실수값을 분수값으로 내림한 포트폴리오 가치를 탐색하는 또다른 동적 계획 해법이다. 이를 통하여 $O(m^3 / \varepsilon)$ 시간에 $(1 - \varepsilon)$-근사해를 출력하는 FPTAS를 얻는다. 넷째는 가상 인스턴스를 푸는 성능이 좋은 모의 담금질(simulated annealing) 기반 휴리스틱 알고리즘이다.

\ref{numericalexperiments}절에서 위에서 도출한 타당성과 계산 복잡성을 확인하는 계산 실험 결과를 정리한다. 모의 담금질 휴리스틱의 정확도도 탐구한다.

결론에서, 해법을 개선할 수 있는 요소를 언급하고, 실제 지원 과정의 특징을 묘사하도록 원래 모형을 확장할 수 있는 기술 몇 가지 소개한다.
\fi

%\pagebreak
\ifen \section{Notation and preliminary results} \else \section{표기법과 예비 결과}\fi\label{preliminaries}
\ifen Before discussing the solution algorithms, we will introduce additional notation and a few preliminary results. For the remainder of the paper, unless otherwise noted, we assume with trivial loss of generality that each $g_j \leq H$, $\sum g_j > H$, each $f_j \in (0, 1]$, and $t_0 < t_1 \leq \cdots \leq t_m$. %Below, we will show how to transform an arbitrary instance so that $t_0 = 0$, in which case each $t_j > 0$. 
\else
해법은 의논하기 전, 본 절에서 추가적인 표기법은 명시하고 몇 가지 유용한 예비 결과를 정리하고자 한다. 남은 논문에서 다른 언급이 없으면, 일반성을 잃지 않고 각 $g_j \leq H$, $\sum g_j > H$, 각 $f_j \in (0, 1]$, 그리고 $t_0 < t_1 \leq \cdots \leq t_m$이라고 가정한다. 뒤에서 임의의 인스턴스를 $t_0 = 0$이 되도록 변환하는 방법을 제시한다. % 이를 적용하면 $t_j > 0$인 가정을 세울 수 있다. 
\fi

\ifen \subsection{Refining the objective function} \else \subsection{목적함수의 정제} \fi

\ifen
We refer to the set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies as her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}. %Let $x$ denote the same encoded as a binary vector, where $x_j=1 \iff j\in \mathcal{X}$ for $j=1\dots m$.
\else
학생이 지원하는 학교 집합 $\mathcal{X} \subseteq \mathcal{C}$를 `지원 포트폴리오'라고 부르자. $\mathcal{X}$에서 학생이 받는 기대 효용을 포트폴리오의 `가치'라고 한다.
\fi 
\begin{definition}[\ifen Portfolio valuation function\else 포트폴리오 가치 함수\fi]
$v(\mathcal{X}) =  \operatorname{E}\left[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\right]$.
\end{definition}
\ifen
\noindent It is helpful to define the random variable $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$ as the utility achieved by the schools in the portfolio, so that when $t_0 = 0$, $v(\mathcal{X}) = \operatorname{E}[X]$. Similar pairs of variables with script and italic names such as $\mathcal{Y}_h$ and $Y_h$ will carry an analogous meaning.
\else
\noindent 편의상 포트폴리오에 속한 학교가 실제로 이루는 효용을 확률 변수 $X  = \max\{ t_j Z_j : j \in \mathcal{X}\}$로 정의한다. 그러면 $t_0 = 0$이면 $v(\mathcal{X}) = \operatorname{E}[X]$임을 알 수 있다. 비슷한 식으로 $\mathcal{Y}_h$와 $Y_h$ 같은 스크립트체와 이탈릭체로 표기한 변수 쌍은 유사한 의미를 가진다.
\fi

\ifen
Given an application portfolio, let $p_j(\mathcal{X})$ denote the probability that the student attends $c_j$. This occurs if and only if she \emph{applies} to $c_j$, is \emph{admitted} to $c_j$, and is \emph{rejected} from any school she prefers to $c_j$; that is, any school with higher index. Hence, for $j= 0\dots m$,
\else
주어진 지원 포트폴리오에 대해 학생이 $c_j$로 진학할 확률을 $p_j(\mathcal{X})$라고 하자. 이 사건이 발생하는 필요충분 조건은 $c_j$에 지원하고, $c_j$에서 합격하고, $c_j$보다 선호하는 (즉, 지표가 더 높은) 학교에서 불합격는 것이다. 따라서  $j= 0\dots m$에 대해
\fi
\begin{align}
p_j(\mathcal{X}) &= 
\begin{cases}
\displaystyle f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}), \quad & j \in \{0\}\cup\mathcal{X}\\
0, \quad & \text{\ifen otherwise\else 그러지 않은 경우.\fi}
\end{cases} 
\end{align}
\ifen
where the empty product equals one and $f_0= 1$. The following proposition follows immediately.
\else
단, 공집합의 곱은 1이면 $f_0= 1$이다. 다음 제의는 직설적인 결과다.
\fi

\begin{proposition}[\ifen Closed form of portfolio valuation function\else 포트폴리오 가치 함수의 다른 식\fi]
\begin{align}
v(\mathcal{X}) &= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
%\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
\end{align}
\end{proposition}
%\begin{proof}Computing $v(\mathcal{X}) = \sum_{j=0}^m  t_j p_j(\mathcal{X})$ yields \eqref{closedformportfoliovaluationX}. Next, because $1 - f_j x_j = 1$ if $x_j = 0$, we may define $p_j$ equivalently as $p_j(x) = x_j  f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’})$ to obtain \eqref{closedformportfoliovaluationx}. 
%\end{proof}

\ifen
Next, we show that without loss of generality, we may assume that $t_0 = 0$.
\else
그다음에, 일반성을 제한하지 않고 $t_0 = 0$이라고 가정할 수 있음을 보이자.
\fi
\begin{lemma} \label{assumetzerozero}
\ifen 
For some $\gamma \leq t_0$, let $\bar t_j = t_j - \gamma$ for $j = 0 \dots m$. Then $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$ regardless of $\mathcal{X}$. 
\else
어떤 $\gamma \leq t_0$을 선택하고 $j \dots m$에 대해 $\bar t_j = t_j - \gamma$라고 하자. 그러면 $\mathcal{X}$를 막론하고 $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$가 성립한다.
\fi
\end{lemma}
\begin{proof}
\ifen
By definition, $\sum_{j=0}^m p_j(\mathcal{X}) = \sum_{j \in \{0\}\cup\mathcal{X}} p_j(\mathcal{X}) = 1$. Therefore
\else
정의에 따라 $\sum_{j=0}^m p_j(\mathcal{X}) = \sum_{j \in \{0\}\cup\mathcal{X}} p_j(\mathcal{X}) = 1$이다. 그러면 다음 수식으로 증명을 완성할 수 있다.
\fi
\begin{align}
\begin{split}
v(\mathcal{X}; \bar t_j) &= \sum_{j\in \{0\}\cup\mathcal{X}}  \bar t_j p_j(\mathcal{X})
=\sum_{j\in \{0\}\cup\mathcal{X}} (t_j - \gamma) p_j(\mathcal{X}) \\
&=\sum_{j\in \{0\}\cup\mathcal{X}} t_j p_j(\mathcal{X})  - \gamma 
= v(\mathcal{X}; t_j) - \gamma
\end{split} \ifen\else\qedhere\fi
\end{align}
\ifen which completes the proof.\fi
\end{proof}

>>>>

In the present study, we formulate the student's college application decision as the following optimization problem:
\begin{align} \label{headlineproblem}
\begin{split}
\text{maximize}\quad & v(\mathcal{X}) =  \operatorname{E}\Bigl[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~|\mathcal{X}| \leq H
\end{split}
\end{align}
Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $h > 0$ is the number of students the student can apply to; for $j = 1 \dots m$, $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j\geq 0$ is a utility parameter.

Consider a single prospective student in a college market with $m$ colleges, and let each $t_j$-value indicate the utility she associates with attending school $j$, where her utility is $t_0$ if she does not attend college. Let $h$ denote the number of schools to which the student can afford to apply. Lastly, let $f_j$ denote the student's probability of being admitted to school $j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not. It is appropriate to assume that the $Z_j$ are statistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates). Then the student's objective is to maximize the expected utility associated with the best school she is admitted to. Therefore, her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies.

% def greedy, nested
We focus on the special case in which students are constrained in the \emph{number} of schools to which they can apply, and show that the optimal portfolios in this case are nested in the budget constraint, implying the optimality of the greedy algorithm that adds schools one at a time according to how much they increase the objective value.

@@

The objective function is a monotonic submodular function in the sense first described by Nemhauser et al. (1978). They showed that the greedy algorithm is asymptotically $(1 - 1/e)$-optimal for optimizing a monotonic submodular function over a cardinality constraint, and a subsequent result of Nemhauser and Wolsey (1978) proved that this approximation ratio is the best achievable by a polynomial-time algorithm. In the special case of college application, however, we show that the same algorithm is exact.

% define nestedness
The proof of the optimality of the greedy algorithm rests on the nested structure of the portfolios, a property that is of independent interest. As Rozanov and Tamir (2020) articulate, the knowledge that the optima are nested aids not only in computing the optimal solution, but in the implementation thereof under uncertain information. For example, in the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges reward students who tailor their essays to the target school. However, students may not know how many schools they can afford to apply to until late October. The nestedness property---or equivalently, the validity of a greedy algorithm---implies that even in the absence of complete budget information, students can begin to carry out the optimal application strategy by writing essays for schools in the order that they enter the optimal portfolio.
% Greedy algorithms produce a \emph{nested} family of solutions parameterized by the budget $H$: If $H \leq H'$, then the greedy solution for budget $H$ is a subset of the greedy solution for budget $H'$.

@@

% talk about DA

% pandora's box problem

@@

\section{Preliminaries} \label{sectionModel}

For the remainder of the paper, unless otherwise noted, we assume with trivial loss of generality that each $f_j \in (0, 1]$ and $< t_1 \leq \cdots \leq t_m$. Unless otherwise noted, we assume $t_0 = 0$, an assumption justified presently. Unless otherwise noted, we assume that $t_0 = 0$, a restriction that we justify presently.

% Lots of overlap here with paper_ellis; edit there first and then import

\subsection{The objective function} \label{sectionObjective}

First we derive a closed-form expression for the objective function of \eqref{headlineproblem}.

We refer to the set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies as her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}. Given an application portfolio, let $p_j(\mathcal{X})$ denote the probability that the student attends school $j$. This occurs if and only if she \emph{applies} to school $j$, is \emph{admitted} to school $j$, and is \emph{rejected} from any school she prefers to $j$; that is, any school with higher index. Hence, for $j= 0\dots m$,
\begin{align}
p_j(\mathcal{X}) &= 
\begin{cases}
\displaystyle f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}), \quad & j \in \{0\}\cup\mathcal{X}\\
0, \quad & \text{otherwise}
\end{cases} 
\end{align}
The following proposition follows by computing $v(\mathcal{X}) = \sum_{j=0}^m  t_j p_j(\mathcal{X})$.
\begin{proposition}[Closed form of portfolio valuation function]
\begin{align}
v(\mathcal{X}) &= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
%\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
\end{align}
\end{proposition}
%\begin{proof}Computing $v(\mathcal{X}) = \sum_{j=0}^m  t_j p_j(\mathcal{X})$ yields \eqref{closedformportfoliovaluationX}. Next, because $1 - f_j x_j = 1$ if $x_j = 0$, we may define $p_j$ equivalently as $p_j(x) = x_j  f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’})$ to obtain \eqref{closedformportfoliovaluationx}. 
%\end{proof}

Next, we show that without loss of generality, we may assume that $t_0 = 0$.

\begin{lemma} \label{assumetzerozero}
For some $\gamma \leq t_0$, let $\bar t_j = t_j - \gamma$ for $j = 0 \dots m$. Then $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$ for any $\mathcal{X} \subseteq \mathcal{C}$. 
\end{lemma}
\begin{proof}
By definition, $\sum_{j=0}^m p_j(\mathcal{X}) = \sum_{j \in \{0\}\cup\mathcal{X}} p_j(\mathcal{X}) = 1$. Therefore
\begin{align}
\begin{split}
v(\mathcal{X}; \bar t_j) &= \sum_{j\in \{0\}\cup\mathcal{X}}  \bar t_j p_j(\mathcal{X})
=\sum_{j\in \{0\}\cup\mathcal{X}} (t_j - \gamma) p_j(\mathcal{X}) \\
&=\sum_{j\in \{0\}\cup\mathcal{X}} t_j p_j(\mathcal{X})  - \gamma 
= v(\mathcal{X}; t_j) - \gamma
\end{split} 
\end{align}
which completes the proof.
\end{proof}

\subsection{An elimination technique} \label{eliminationtechniquesection}

Now, we present a variable-elimination technique that will prove useful throughout the paper.\footnote{We thank Yim Seho for pointing out this useful transformation.} Suppose that the student has already resolved to apply to school $k$, and the remainder of her decision consists of determining which \emph{other} schools to apply to. Writing her total application portfolio as $\mathcal{X} = \mathcal{Y} \cup \{k\}$, we devise a function $w(\mathcal{Y})$ that orders portfolios according to how well they ``complement'' the singleton portfolio $\{k\}$. Specifically, the difference between $v(\mathcal{Y} \cup\{k\})$ and $w(\mathcal{Y})$ is the constant $f_k t_k$.

To construct $w(\mathcal{Y})$, let $\tilde t_j$ denote the expected utility the student receives from school $j$ \emph{given} that she has been admitted to school $j$ and applied to school $k$. For $j < k$ (including $j = 0$), this is $\tilde t_j = (1- f_k) t_j + f_k t_k$; for $j > k $, this is $\tilde t_j = t_j$. This means that 
\begin{equation}\label{Vyastildet}
v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
The transformation to $\tilde t$ does not change the order of the $t_j$-values. Therefore, the expression on the right side of \eqref{Vyastildet} is itself a portfolio valuation function. In the corresponding market, $t$ is replaced by $\tilde t$ and $\mathcal{C}$ is replaced by $\mathcal{C}\setminus\{k\}$. To restore our convention that $t_0 = 0$, we obtain $w(\mathcal{Y})$ by taking $\bar t_j = \tilde t_j - \tilde t_0$ for all $j \neq k$ and letting
\begin{equation}  \label{wYvXminusconst}
w(\mathcal{Y})
= \sum_{j \in \{0\} \cup \mathcal{Y}} \bar t_j p_j(\mathcal{Y})
= \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
= v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
where the second equality follows from Lemma \ref{assumetzerozero}. The validity of this transformation is summarized in the following theorem, where we write $v(\mathcal{X}; \bar t)$ instead of $w(\mathcal{Y})$ to emphasize that $w(\mathcal{Y})$ is, in form, a portfolio valuation function.

\begin{lemma}[Eliminate school $k$] \label{eliminationtheorem}
For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
\begin{align}\label{howtotransformtj}
\bar t_j = 
\begin{cases}
(1 - f_k) t_j, \quad & t_j \leq t_k \\
t_j - f_k t_k, \quad& t_j > t_k.
\end{cases}
\end{align}
\end{lemma}

\begin{proof}
It is easy to verify that \eqref{howtotransformtj} is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) discussed above.
\end{proof}

%\noindent The transformation \eqref{howtotransformtj} can be applied iteratively to accommodate the case where the student has already resolved to apply to multiple schools.

\subsection{Submodularity of the objective}

Now, we show that the portfolio valuation function is submodular. This result is primarily of taxonomical interest and may be safely skipped, as our subsequent results do not rely on submodular analysis.

\begin{definition}[Submodular set function]
Given a ground set $\mathcal{C}$ and function $v : 2^{\mathcal{C}} \mapsto \mathbb{R}$, $v(\mathcal{X})$ is called a \emph{submodular set function} if and only if $v(\mathcal{X}) + v(\mathcal{Y}) \geq v(\mathcal{X}\cup\mathcal{Y}) + v(\mathcal{X}\cap\mathcal{Y})$
for all $\mathcal{X}, \mathcal{Y} \subseteq \mathcal{C}$. Furthermore, if $ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \geq 0$ for all $\mathcal{X} \subset \mathcal{C}$ and $k \in \mathcal{C} \setminus \mathcal{X}$, $v(\mathcal{X})$ is said to be a \emph{nondecreasing} submodular set function.
\end{definition}

\begin{theorem}
The college application portfolio valuation function
$v(\mathcal{X})$ % = \sum_{j\in\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\]
is a nondecreasing submodular set function.
\end{theorem}

\begin{proof}
It is self-evident that $v(\mathcal{X})$ is nondecreasing. To establish its submodularity, we apply proposition 2.1.iii of Nemhauser and Wolsey (1978) and show that
\begin{equation}\label{nemhauseriii}
v(\mathcal{X} \cup \{j\}) - v(\mathcal{X}) \geq 
v(\mathcal{X} \cup \{j, k\}) - v(\mathcal{X} \cup \{k\})
\end{equation}
for $\mathcal{X} \subset \mathcal{C}$ and $j \neq k \in \mathcal{C} \setminus \mathcal{X}$. By Lemma \ref{eliminationtheorem}, we can repeatedly eliminate the schools in $\mathcal{X}$ according to \eqref{howtotransformtj} to obtain a portfolio valuation function $w(\mathcal{Y})$ with parameter $\bar t$ such that $w(\mathcal{Y}) = v(\mathcal{X} \cup \mathcal{Y}) + \text{const.}$ for any $\mathcal{Y} \subseteq \mathcal{C} \setminus \mathcal{X}$. Therefore, \eqref{nemhauseriii} is equivalent to
\begin{align}
& w(\{j\}) - w(\varnothing) \geq w(\{j, k\}) - w(\{k\}) \\
\iff \qquad &w(\{j\})  +  w(\{k\})  \geq w(\{j, k\})  \\
\iff \qquad &\operatorname{E}[\,\bar t_j Z_j\,] + \operatorname{E}[\,\bar t_k Z_k\,] 
\geq \operatorname{E}\bigl[\max\{ \bar t_j Z_j, \bar t_k Z_k \} \bigr]
\end{align}
which is plainly true. 
\end{proof}

@@

\subsection{The nestedness property} 
The optimality of the greedy algorithm for the college application problem rests on the fact the the optimal solution possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset.

\begin{theorem}[Nestedness of optimal application portfolios] \label{nestedapplication}
There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation

\begin{equation}
\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m
\end{equation}
such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.
\end{theorem}

\begin{proof}
By induction on $h$. Applying Lemma \ref{assumetzerozero}, we assume that $t_0 = 0$.

(Base case.) First, we will show that $\mathcal{X}_1 \subset \mathcal{X}_2$. To get a contradiction, suppose that the optima are $\mathcal{X}_1 = \{j\}$ and $\mathcal{X}_2 = \{k, l\}$, where we may assume that $t_k \leq t_l$. Optimality requires that

\begin{equation}v(\mathcal{X}_1 )  = f_j t_j > v(\{k\}) = f_k t_k\end{equation}
and
\begin{align}
\begin{split}
v(\mathcal{X}_2) =  f_k (1- f_l) t_k + f_l t_l &> v(\{j, l\}) \\
& = f_j (1- f_l) t_j + (1- f_j) f_l t_l + f_j f_l \max\{t_j, t_l\} \\
&\geq  f_j (1- f_l) t_j + (1- f_j) f_l t_l + f_j f_l  t_l \\
&= f_j (1- f_l) t_j + f_l t_l  \\
&\geq f_k (1- f_l) t_k + f_l t_l  = v(\mathcal{X}_2)
\end{split}
\end{align}
which is a contradiction.

(Inductive step.) Assume that $\mathcal{X}_1 \subset \cdots \subset \mathcal{X}_h$, and we will show $\mathcal{X}_h \subset \mathcal{X}_{h+1}$. Let $k = \argmax\{ t_k: k \in \mathcal{X}_{h+1}\}$ and write $\mathcal{X}_{h+1} = \mathcal{Y}_{h} \cup \{k\}$.

Suppose $k \notin \mathcal{X}_h$. To get a contradiction, suppose that $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$ and  $v(\mathcal{X}_{h+1}) > v(\mathcal{X}_h \cup \{k\})$. Then
\begin{align}
\begin{split}
v(\mathcal{X}_{h+1})&= v(\mathcal{Y}_{h} \cup \{k\}) \\
&= (1 - f_k) v(\mathcal{Y}_h) + f_k t_k \\
&\leq (1 - f_k) v(\mathcal{X}_h) + f_k \operatorname{E}\bigl[ \max\{t_k, X_h\}\bigr]\\
&=  v(\mathcal{X}_h\cup \{k\})
\end{split}
\end{align}
is a contradiction.

Now suppose that $k \in \mathcal{X}_h$. We can write $\mathcal{X}_h = \mathcal{Y}_{h-1} \cup \{k\}$, where $ \mathcal{Y}_{h-1}$ is some portfolio of size $h-1$. It suffices to show that $ \mathcal{Y}_{h-1} \subset \mathcal{Y}_h$. By definition, $\mathcal{Y}_{h-1}$ (respectively, $\mathcal{Y}_{h}$) maximizes the function $v(\mathcal{Y}\cup\{k\})$ over portfolios of size $h-1$ (respectively, $h$) that do not include $k$. That is, $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$ are the optimal complements to the singleton portfolio $\{k\}$.

Applying Lemma \ref{eliminationtheorem}, we eliminate school $k$, transform the remaining $t_j$-values to $\bar t_j$ according to \eqref{howtotransformtj}, and obtain a function $w(\mathcal{Y}) = v(\mathcal{Y} \cup \{k\}) - f_k t_k$ that grades portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$. Since $w(\mathcal{Y})$ is itself a portfolio valuation funtion and $\bar t_0 = 0$, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.
\end{proof}

\subsection{Quadratic-time solution}

Applying Theorem \ref{nestedapplication} yields an efficient greedy algorithm for the optimal portfolio: Start with the empty set and add schools one at a time, maximizing $v(\mathcal{X}\cup \{k\})$ at each addition. Sorting $t$ is  $O(m \log m)$.  At each of the $h$ iterations, there are $O(m)$ candidates for $k$, and computing $v(\mathcal{X}\cup \{k\})$ is $O(h)$ using \eqref{closedformportfoliovaluationX}; therefore, the time complexity of this algorithm is $O(h^2 m + m \log m)$.

We reduce the computation time to $O(hm)$ by taking advantage of the transformation from Lemma \ref{eliminationtheorem}. Once school $k$ is added to $\mathcal{X}$, we eliminate it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to \eqref{howtotransformtj}. Now, the \emph{next} school added must be the optimal singleton portfolio in the modified market. But the optimal singleton portfolio consists simply of the school with the highest value of $f_j \bar t_j$. Therefore, by updating the $t_j$-values at each iteration according to \eqref{howtotransformtj}, we eliminate the need to compute $v(\mathcal{X})$ entirely. Moreover, this algorithm does not require the schools to be indexed in ascending order by $t_j$, which removes the $O(m\log m)$ sorting cost.

\begin{algorithm}[h] 
%\DontPrintSemicolon
\caption{Optimal portfolio algorithm for the college application problem.} \label{algorithmforlargeh}
\KwIn{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application limit $h \leq m$.}
$\mathcal{C} \gets \{1 \dots m\}$\;
$\mathtt{X, V} \gets $ empty $h$-arrays\;
\For{$i=1\dots h$}
{
    $k \gets \argmax_{j \in \mathcal{C}}\{f_j t_j\}$\;
    $\mathcal{C} \gets \mathcal{C} \setminus \{k\}$\;
    $\mathtt{X}[i] \gets k$\;
     \lIfElse{$i=1$}{$\mathtt{V}[i] \gets f_k t_k$}
     {$\mathtt{V}[i] \gets \mathtt{V}[i-1] + f_k t_k$}
    \For{$j \in \mathcal{C}$}
	{
	\lIfElse{$t_j \leq t_k$}{$t_j \gets (1 -  f_k) t_j $}{$t_j \gets  t_j -  f_k t_k$}
	}
}
\Return{$\mathtt{X, V}$}
\end{algorithm}

\begin{theorem}[Validity of Algorithm \ref{algorithmforlargeh}] \label{validityofalmaalgorithm}
Algorithm \ref{algorithmforlargeh} produces an optimal application portfolio for the cardinality-constrained college application problem in $O(h m)$ time.
\end{theorem}

\begin{proof}
Optimality follows from the proof of Theorem \ref{nestedapplication}. At each of the $h$ iterations of the main loop, finding the top school costs $O(m)$, and the $t_j$-values of the remaining $O(m)$ schools are each updated in unit time. Therefore, the overall time complexity is $O(h m)$.
\end{proof}

By running the algorithm with $h = m$, one can obtain the optimal portfolio for \emph{all} $h$ by taking $\mathcal{X}_h = \{ \mathtt{X}[1], \dots, \mathtt{X}[h]$. Thus, Fu (2014)'s variant of the college application problem 
\[ \text{maximize} \quad v(\mathcal{X}) - c(|\mathcal{X}|)\]
(where $c(|\mathcal{X}|$ is an arbitrary increasing cost function) can be solved in $O(m^2)$-time by a trivial modification of Algorithm \ref{algorithmforlargeh}: Simply compute
\[h^* = \argmax_{h = 1 \dots m} \bigl\{\mathtt{V}[h] - c(h)\bigr\}\] and return $\mathcal{X}_{h^*}$.

@@

\parskip 0em
\leftskip 2em
\parindent -2em

\indent\indent Acharya, Mohan S., Asfia Armaan, and Aneeta S. Antony. 2019. ``A Comparison of Regression Models for Prediction of Graduate Admissions.'' In \emph{Second International Conference on Computational Intelligence in Data Science.} \url{https://doi.org/10.1109/ICCIDS.2019.8862140}.

@@
